---
title: "BIBEVegetation_Analyses_2020"
author: "Julie Coffey"
date: "April 26, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("~/Big-Bend-Data")
require(vegan)#dist matrices, NMDS stuff
require(ggplot2)#plotting
require(tidyverse)#data shaping
require(dplyr)# data manip
require(reshape)
```

## BIBE Vegetation Analyses

This script takes vegetation & remotely-sensed site characteristics and looks for differences between and among sites in three restoration groupings via ordination anova and anosim.

```{r read in data}
env <- read.csv("Environmental_df2020.csv",sep = ",")
head(env)
dim(env) #67  sites by 7 vars
```

## Create a distance matrix. 

Here we will first standardize since our variables are on different scales, then create the matrix using Euclidean as a distance measure.

```{r Distance matrix, echo=FALSE}
rownames(env) <- env$site
s.env <- scale(env[,2:6])
head(s.env)

d <- dist(s.env, method ="euclidean")
head(d)
d
```

## Ordinate!
Okay, now let's visualize these distances with an ordination 

```{r Ordinaton}
vord <- metaMDS(comm = d,k = 3,trymax = 200,autotransform = F, wascores = T)
stressplot(vord)
vord  #stress = 0.04087459

#Create a dataframe with the results for displaying
scors <- as.data.frame(scores(vord, display = "sites")) #save NMDS results into dataframe
scors <- cbind(scors, Management = env$Restoration_group) #add grouping variable "Management" to dataframe
scors <- cbind(scors, Site = rownames(scors)) #add site names as variable if you want to display on plot
head(scors)

```
## Plot our ordination

```{r plot}
veg.mds<- ggplot(scors, aes(x=NMDS1, y=NMDS2))+ 
  geom_point(aes(NMDS1, NMDS2, shape = factor(Management), colour = factor(Management)), size = 3)+   
  coord_fixed()+
  scale_colour_grey()+
  theme(panel.background = element_rect(fill = NA, colour = "black", size = 1, linetype = "solid"))+
  labs(colour = "Restoration Group", shape = "Restoration Group")+ 
  theme(legend.position = "right", legend.text = element_text(size = 12), legend.title = element_text(size = 12), axis.text = element_text(size = 10)) 

veg.mds

#ggsave(veg.mds,filename = "Vegetation_ord_2020.png",device = "png",width = 8,height = 5,units = "in",dpi = 320)
       
```

